Stagnating State Funding for Higher Education and its Effect on Faculty Composition at US Universities

### 2023-05-20 text/paper.tex
Adjusting writing in TeX document.  2nd revision unfinished.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title page
% Author
\author{Senan Hogan-Hennessy \\ \vspace{0.1cm}
    Economics Department, Cornell University\thanks{
        I thank Professor Evan Riehl for guidance on this project, Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, and Ryan Dycus for various comments on the project's direction, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.}
}
% Title
\title{Changes in Faculty Composition and Stagnating State Support for Higher Education
    }
\date{This version: \today\thanks{
    This project's Github repository, with materials and data (where publicly available) for replication, is available at 
    \url{https://github.com/shoganhennessy/state-faculty-composition}.
    Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
    } \\ \vspace{0.1cm}
    \href{https://github.com/shoganhennessy/state-faculty-composition/blob/main/state-faculty-composition-2023.pdf}{Newest version available here.}
}
\rhead{\today}
\lhead{Faculty Composition and State Support for Higher Education.}
% Begin
\begin{document}
\maketitle
\thispagestyle{empty}
% Abstract
\begin{abstract}
    \input{sections/00-abstract.tex}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Starting the paper
\newpage
\setcounter{page}{1}
\doublespacing
% Introduction section
\input{sections/01-introduction.tex}
% Data section
\input{sections/02-data.tex}
% Conceptual Framework
\input{sections/03-concepts.tex}
% Empirical methods
\input{sections/04-empirical.tex}
% Empirical results
\input{sections/05-results.tex}
% Dicussion
\input{sections/06-discussion.tex}
% Conclusion
\input{sections/07-conclusion.tex}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title page
% Author
\author{Senan Hogan-Hennessy\thanks{
        I thank Professor Evan Riehl for his guidance on this project, without which this work would not have possible.
        Additionally, I thank Michael Lovenheim, Ronald Ehrenberg, Douglas Miller, Francine Blau, Levon Barseghyan, Jake Meyer, and Ryan Dycus for various comments, as well as seminar participants at Cornell University (Autumn 2022) for helpful discussion.} \\
    \vspace{0.1cm}
    Economics Department, Cornell University
}
% Title
\title{Stagnating State Funding for Higher Education and its Effect on Faculty Composition at US Universities
    }
\date{This version: \today\thanks{
    This project's Github repository, with materials and anonymised data for replication, is available at 
    \url{https://github.com/shoganhennessy/state-faculty-composition}.
    Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.
    } \\ \vspace{0.1cm}
    \href{https://github.com/shoganhennessy/state-faculty-composition/blob/main/state-faculty-composition-2023.pdf}{Newest version available here.}
}
\rhead{\today}
\lhead{Stagnating State Funding and its Effect on Faculty Composition.}
% Begin
\begin{document}
\maketitle
\thispagestyle{empty}
% Abstract
\begin{abstract}
    \noindent
    \input{sections/00-abstract.tex}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Starting the paper
\newpage
\setcounter{page}{1}
\doublespacing
% Introduction section
\noindent
\input{sections/01-introduction.tex}
% Data section
\input{sections/02-data.tex}
% Conceptual Framework
\input{sections/03-concepts.tex}
% Empirical methods
\input{sections/04-empirical.tex}
% Empirical results
\input{sections/05-results.tex}
% Dicussion
\input{sections/06-discussion.tex}
% Conclusion
\input{sections/07-conclusion.tex}

### 2023-05-20 text/sections/00-abstract.tex
Adjusting writing in TeX document.  2nd revision unfinished.

\noindent
State support for higher education per-student has stagnated while universities have substituted towards contingent lecturers and away from tenure-track/tenured professors.
This analysis relates these changes in the faculty composition at public universities to the fall in state funding per student for higher education, using a shift-share approach to identify changes in a state funding among public universities.
A decrease in state funding of 10\%, via a shock to state funding, decreases the number of assistant professors per student at a public university by 1.4\% and full professors by 1.2\%, yet increases the number of lecturers per student by 4.4\%.
Local projection estimates show that these effects linger for up to three years after the initial funding shock.
Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by impacting the hiring of new professors at public universities impacted by the funding shocks.
These results show the long-term effects of stagnating state support for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

>>>>

While states have limited their funding for higher education for the last forty years, US universities have systematically substituted away from tenure-track/tenured professors towards contingent lecturers.
This analysis relates the changes in the faculty composition at US public universities to the fall in state funding for higher education, using a shift-share approach to isolate the causal effects of changes in state funding.
A $-10$\% fall in a university's funding, via a state funding shock, decreases the number of assistant professors per student by $-1.4$\% and full professors by $-1.2$\%, yet increases the number of lecturers per student by 4.4\%; long-run estimates show the effects linger for multiple years after the initial funding shock.
Analysis of all the professors at Illinois public universities 2011-2021 shows that incumbent professors are not affected by the changes in state funding, and suggestive evidence shows that these changes arose by limiting the hiring of new professors at public universities to replace leaving tenure-track/tenured professors.
These results show the institutional effects of stagnating state funding for higher education, and raise questions about the direction that public education heads as these financial headwinds show no sign of dissipating.

### 2023-05-20 text/sections/01-introduction.tex
Adjusting writing in TeX document.  2nd revision unfinished.

\noindent
Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
This decline has been shown to lead to worse education and later-life outcomes for students, yet it is not clear what mechanisms drive these effects, or how public universities are affected as institutions.
I show that four-year, degree-granting public universities have experienced falling state funding per student, and this affected their composition of faculty.
Falls in public university revenues from the state, instrumented by state-level funding shocks, lead to a fall in the number of tenure-track or tenured professors per student within a university, and an increase in the number of lecturers.
US private universities were not exposed to such financial constraints during this same time period, and do not exhibit the substitution away from tenure-track and tenured professors, so that the stagnating state support for public universities has implications for the wider structure of higher education instruction and research in the US.

US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 per enrolled student in 1990, yet only \$8,300 per enrolled student in 2021.
This fall is driven by stagnating funding provided by state governments, while enrolment among public universities rose by 46\% over the same time period.
At the same time, the number of professors per student at private universities stayed relatively stable at 0.045, yet fell from 0.04 to around 0.35 professors per student at public universities.
This change was driven by a fall of over 20\% in the count of associate and full professors per student at public universities.

To identify changes in state funding for higher education, I employ a shift-share instrument for state funding of public universities, which exploits reliance on state funding and yearly shocks to the amount an entire state funds higher education \citep{NBERw23736,NBERw27885}.
State appropriations for public universities are plausibly endogenous to many outcomes: state governments decide on yearly budgets for their higher education sector, and this process can be influenced by local financial conditions, or changing state priorities.
The shift-share instrument interacts reliance on state funding in a base period with the yearly total appropriations per student in the entire state, to exploit both changes in support for higher education and how much each university historically has relied on that support.

Negative changes in state funding affects faculty composition at public universities, away from tenured and tenure-track professors towards non-tenure track lecturer positions.
A reduction of 10\% in state funding, via a shock to state appropriations, leads to a fall in 1.4\% in the number of assistant professors per student within a university, 1.2\% for full (tenured) professors, and an increase in 4.4\% the number of lecturers per student.
Local projection estimates show that these effects linger for three to four years after the initial funding shock.
Over the same time period, state funding fell by around 35\%, while the count of professors per student fell by 9\%, so that these results show that falls in state funding explain around a third of the observed shift away from tenure-track and tenure professors towards contingent faculty.
Additionally, I use individual level data on professors at Illinois public universities over 2010-2021 to investigate whether the shocks to state funding affected individual professors.
Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
Yet, the hiring rate of tenure-track professors declined over this time period for all public universities.
This shows that changes in faculty composition came about by hiring reductions of new professors at public universities.

@@

A number of studies have shown that stagnating state funding for higher education negatively impacts student outcomes.
\cite{NBERw23736} show that increases in public university spending (via funding shocks) increases enrolment and degree completion among students,\footnote{
    \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
}
that high school students exposed to similar public school funding shocks are less likely to attend college \citep{jackson2021school}, and that these shocks induce public universities to rely more on tuition as a source of revenue \citep{bound2019public}.
\cite{NBERw27885} use the same methodology to show that increases in state appropriations lead to degree lower completion time\footnote{
    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower higher education completion rate at public universities, 
} and later-life debt for students.

Faculty are a core component of the US higher education and research-innovation university system, yet their composition within universities, and its relationship with higher education finances has so far not been studied.
The closest previous examples study  how university endowments react to negative endowment shocks \citep{brown2014endowment} and the 2008 recession \citep{turner2014impact}, and multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.

The number of professors, both in absolute terms and relative to number of students, may be particularly important for quality of instruction and thus educational outcomes.
Small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
Regarding faculty composition, the literature is not decided on whether contingent lecturers lead to better or worse education outcomes relative to tenured faculty.
Adjunct professors are shown to better undergraduate teachers at highly selective universities \citep{bettinger2010does,figlio2015tenure}, yet universities who use a greater share of lecturers have lower graduation rates \citep{ehrenberg2005tenured}.
Notably, students taught by part-time adjuncts lecturers have worse student outcomes than full-time professors at a less-selective university \citep{zhu2021limited}, but this difference disappears when individual lecturers gain a full-time teaching position at the university. 
My approach focuses on count of professors per student at universities, and considers faculty count and composition change as possible mechanisms for how stagnating state support for higher education negatively impacts student outcomes \citep{NBERw23736,NBERw27885}.

>>>>

Public universities educate the majority of higher education students in the US, yet have experienced a secular decline in state funding (per student) the last three decades.
At the same time, public university student enrollment and employment of contingenet lecturers dramtically increased at public universities.
I show that four-year, degree-granting public universities have experienced a systematic fall in state funding per student, and this fall led to a substitution away from tenure-track/tenured professors towards contingent lecturers.
Falls in state funding for public universities, instrumented by state-level shocks to state-wide higher education, led to a fall in the number of tenure-track/tenured professors per student within a university, and an increase in the number of lecturers.
Private universities were not exposed to similar financial constraints during this same time period, and do not exhibit the substitution away from tenure-track/tenured professors, so that the stagnating state support for public universities has implications for the wider structure of US higher education and research.

US universities are widely considered the highest performing in the world, yet there are consequential differences between its universities that operate in the private sector and those established by state governments.
Public universities are subject to numerous state-level administrative laws, and rely on their state governments for funding: an average public university received around \$11,600 of state funding per student in 1990, and only \$8,300 in 2021.
This fall is driven by a stagnation in the absolute level of state funding for higher education in most states, combined with a large increase of 46\% in student enrolment at public universities over the same time period.
Similarly, the number of lecturers per student over doubled, while the number of tenure-track/tenured professors per students fell by 23\% from 1990 to 2021.

I employ a shift-share instrument to identify changes in state funding for higher education, which exploits how universities differ in their financial reliance on state funding, combined with yearly shocks to the amount an entire state funds higher education --- following \cite{NBERw23736,NBERw27885}.
State governments decide how to fund higher education, and everything else, by a complicated process which can be influenced by local economic conditions, changing state priorities, or even lobbying from the state universities themselves.
As such, it is necessary to employ an identification strategy, such as the shift-share instrument, to account for endogeneity in universities' state funding.
The shift-share instrument interacts reliance on state funding in a base period with the entire state' total funding per student for each year, to exploit how much each university historically relied on state funding and yearly changes in state funding for higher education.

Falls in state funding affects faculty composition at public universities, who substitute away from tenured and tenure-track professors towards contingent lecturers.
A fall of $-10$\% in state funding leads to a fall of $-1.4$\% in the number of assistant professors per student at a university, a fall of $-1.2$\% for tenured professors, and an increase in 4.4\% the number of lecturers per student.
Local projection estimates show that these effects linger for three to four years after the initial funding shock, showing that the effect is not isolated to the immediate year of the first funding shock. 
Over the same time period, state funding fell by around $-$35\%, while the count of professors per student fell by $-$9\%, so that these results show that falls in state funding explain around a third of the substitution away from tenure-track and tenure professors towards contingent lecturers.
Additionally, I use individual level data on all Illinois public university professors in 2010-2021 to investigate whether the shocks to state funding affected individual professors.
Incumbent professors were not meaningfully affected, in terms of total salary, promotion rate, or rate of leaving the Illinois public university system.
Yet, the hiring rate for professors declined over this time period for all public universities, and particularly for universities with the greatest falls in state funding.
This implies that faculty composition change arose by limiting the hiring of new professors; public universities increasingly hired contingent lecturers, and increasingly did not replace their retiring (or leaving) professors.

@@

A number of studies have shown that stagnating state funding for higher education negatively impacts student outcomes.
\cite{NBERw23736} show that increases in public university spending (via funding shocks) increases enrolment and degree completion among students,\footnote{
    \cite{miller2022making} further analyse the effects of falls in university revenues, by finding that reductions in public universities tuition prices (suggestively motivated by falling state funding) leads to reductions in provided financial aid.
}
that high school students exposed to similar public school funding shocks are less likely to attend college \citep{jackson2021school}, and that these shocks induce public universities to rely more on tuition as a source of revenue \citep{bound2019public}.
\cite{NBERw27885} use the same methodology to show that increases in state funding lead to degree lower completion time and later-life debt for students.\footnote{
    \cite{bound2007cohort} use similar variation in state funding per student to show that lower funding leads to lower completion rates at public universities.
}

Faculty composition is an interesting outcome on its own, but its relationship with higher education finances has so far not been studied.
The closest previous examples study  how university endowments react to negative endowment shocks \citep{brown2014endowment} and the 2008 recession \citep{turner2014impact}, and multiple theoretically model university decision-making and faculty hiring (see e.g., \citealt{abe2015implications,johnson2009jep,NBERc13879}).
%Lastly, instruction costs within individual universities vary substantially across subject areas, primarily thanks to department differences in class size and faculty salaries \citep{hemelt2021math}.
The number of professors per student may be particularly important for quality of instruction and thus educational outcomes.
Small class sizes in secondary schools lead to increases in test scores \citep{angrist1999using}, and UK university students perform worse academically in particularly large classes \citep{bandiera2010heterogeneous}.
Though, the economics literature is not decided on whether contingent lecturers lead to better or worse outcomes in higher education.
Adjunct professors are generally more effective teachers at highly selective universities \citep{bettinger2010does,figlio2015tenure}, yet students taught by part-time lecturers have worse student outcomes \citep{zhu2021limited}, and universities who use a greater share of lecturers have lower graduation rates \citep{ehrenberg2005tenured}.

While it is not immediately clear whether a systematic substitution towards lecturers away from professors reduces teaching effectiveness, it is clear that stagnating state funding has negatively impacted student outcomes \citep{NBERw23736,NBERw27885}.
This analysis focuses on a university's count of faculty per student, and considers faculty composition change as one possible mechanism for how stagnating state funding for higher education has negatively affected students at public universities for the last 40 years.

### 2023-05-20 text/sections/02-data.tex
Adjusting writing in TeX document.  2nd revision unfinished.

The institutional environment for how states appropriate funding allows for wide variation in state funding for higher education, both between states and across years in each state.
A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
    \cite{NBERw23736} present a full discussion of the decision-making process for state appropriations, drawing on administrative records originally analysed by \cite{parmley2009state}.
}
US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
This process leads to yearly variation in state appropriations not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
State appropriations to higher education are a particularly attractive area of state spending to absorb negative shocks to state finances \citep{delaney2011state}.\footnote{
    \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends, and \autoref{fig:funding} observes these same trends.
}
Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state appropriations vary on multiple dimensions.

Public university revenues on average stagnated across there years 1990--2021.
\autoref{fig:funding} show the trends in revenues for the mean public university for the years 1990--2021.
We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
At the same time, total state appropriations stagnated at around \$100 million per year-university for 1990--2008, falling around 2008 and have not recovered ever since.
While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Mean Total Revenues among Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, millions \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
        \label{fig:mean-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Enrolled Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
        \label{fig:mean-funding-fte}
    \end{subfigure}
    \label{fig:funding}
\end{figure}

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Total Student enrolment, by University Sector, and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total.}
        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
        \label{fig:enrollment-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Mean, University Level.}
        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
        \label{fig:enrollment-mean}
    \end{subfigure}
    \label{fig:enrolment}
\end{figure}

@@

The rate that Illinois funds higher education has stagnated to a similar extent as the average for the entire country.
\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state appropriations and substitution towards tuition revenue.

@@

The decade of 2010--2021 has also seen yearly variation in its revenues, particularly around 2016.
In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
    Fiscal year 2016 refers to June 205 to June 2016, so is the same as the academic year definition.
}
While public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
This variation in public university revenues via the state appropriations channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state appropriations over 2010-2021, of similar order to those for the rest of the country over 1990--2021.

>>>>

The institutional environment for how states appropriate funding allows for wide variation in state funding for higher education, both between states and across years in each state.
A state government plans an annual budget a couple of years ahead of the fiscal year, and the legislature chooses to approve or reject a budget request put forth by the governor's office.\footnote{
    \cite{NBERw23736} present a full discussion of the decision-making process for state appropriations, drawing on administrative records originally analysed by \cite{parmley2009state}.
}
US state governments, by majority, are legally obligated to run a balanced-budget, so that yearly variation in tax revenues (caused by changing economic conditions or otherwise) necessarily affect state expenditures.
This process leads to yearly variation in state funding not seen in other revenues sources (such as federal government appropriations), since US states differ in their support for higher education, and are subject to fiscal constraints.
State funding for higher education are a particularly attractive area of state spending to absorb negative shocks to state finances \citep{delaney2011state}.\footnote{
    \cite{delaney2011state} fully describe the financial environment of state expenditures, and what makes spending on higher education an attractive area for state governments to expand funding during years of higher tax revenues, and retract funding in leaner years.
    An analysis of state expenditures for the years 1980-2004 (overlapping with the sample for this analysis) provides solid evidence for these trends, and \autoref{fig:funding} observes these same trends.
}
Additionally, there is yearly variation in the number of higher education students in each state \citep{turner2014impact}, so that per-student state funding varies on multiple dimensions.

Public university revenues on average stagnated across there years 1990--2021.
\autoref{fig:funding} show the trends in revenues for the mean public university for the years 1990--2021.
We see a rise in total revenues received by public universities (from all sources), and a notable increase in mean tuition revenues from \$48 million per year-university to \$150 million per year-university.
At the same time, total state funding stagnated at around \$100 million per year-university for 1990--2008, falling around 2008 and have not recovered ever since.
While public universities experienced a stagnation in state support, private universities were not exposed to the same constraints, receiving \$37,000 per student in 1990 and \$49,000 in 2021, experiencing no corresponding decline in any specific component.

\begin{figure}[!htbp]
    \centering
    \singlespacing
    \caption{Mean Total Revenues among Public Universities, by Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-total.png}
        \label{fig:mean-funding-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Per Student, \$ 2021 CPI-U.}
        \includegraphics[width=\textwidth]{figures/mean-funding-fte.png}
        \label{fig:mean-funding-fte}
    \end{subfigure}
    \label{fig:funding}
\end{figure}

\begin{figure}[!htbp]
    \centering
    \singlespacing
    \caption{Total Student enrolment, by University Sector and Year.}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Total.}
        \includegraphics[width=\textwidth]{figures/enrollment-total.png}
        \label{fig:enrollment-total}
    \end{subfigure}
    \begin{subfigure}[b]{0.495\textwidth}
        \centering
        \caption{Mean, University Level.}
        \includegraphics[width=\textwidth]{figures/enrollment-mean.png}
        \label{fig:enrollment-mean}
    \end{subfigure}
    \label{fig:enrolment}
\end{figure}

@@

The rate that Illinois funds higher education has stagnated to a similar extent as the average for the entire country.
\autoref{fig:illinois-funding} shows the trends for revenues among the seven Illinois public university campuses, including the falling share of state funding and substitution towards tuition revenue.

@@

The decade of 2010--2021 has also seen yearly variation in its revenues, particularly around 2016.
In the calendar year 2015 partisan disagreements between the democratic legislature and republican governor led to the 2016 fiscal year starting with no state budget.
State agencies, and higher education institutions, employed accounting techniques to continue operating without any resources provided by the state government.\footnote{
    Fiscal year 2016 refers to June 205 to June 2016, so is the same as the academic year definition.
}
While public universities were able to stay open, there were drastic revenue and spending cuts in response to the budget impasse, as it continued through fiscal year 2017, and ended with a new budget restoring funding to state institutions for 2018.
This variation in public university revenues via the state funding channel, stemming entirely from political disagreements --- not from state decisions regarding higher education and its finances \citep{young2020squandered} --- mean that Illinois public universities exhibit sizable changes in their state funding over 2010-2021, of similar order to those for the rest of the country over 1990--2021.
